\documentclass[11pt, oneside]{letter}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\usepackage{amssymb}

%\address{640 Ferry Street \\ Lafayette, Indiana  47901 \\ USA}  
\begin{document}
\begin{letter}
\opening{Dear Search Committee,}
I am writing to apply for JOB. 


I am a Ph.D. candidate at Purdue University, under the direction of D.B. McReynolds. I expect to complete my degree by May of 2023.

My research interests include geometry and topology, spectral theory, number theory, representation theory, harmonic analysis, and interactions thereof. My thesis work concerns an absolute spectral rigidity result, which applies to a large class of arithmetic hyperbolic 2 and 3 manifolds.  One application of this theorem proves that all principal congruence Hurwitz surfaces are spectrally rigid. I have also co-authored a paper in which we construct large collections of algebraic and differential geometric objects which are indistinguishable by means of various invariants. I consider teaching to be an essential role of the mathematician, and am committed to fostering a safe, inclusive, and open environment for students. 

Please find my curriculum vitae and research statement attached. Letters in support of my application, written by D.B. McReynolds, M. Stover, J. LaFont, A. Reid, and J. Chen, will be submitted through the forms on mathjobs.com.  Please contact me for any further information. 

\closing{Sincerely,}
Justin Katz\\
Mathematics Department\\
Purdue University\\
1610 Purdue Mall\\
West Lafayette, IN 47907\\
(219) 210-9452 \\jukatz93@gmail.com\\
https://www.katz9.blog

\end{letter}
\end{document}  