\documentclass[a4paper,oneside,10pt]{letter}
\makeatletter
\renewcommand{\closing}[1]{\par\nobreak\vspace{\parskip}%
  \stopbreaks
  \noindent
  \ifx\@empty\fromaddress\else
  \hspace*{\longindentation}\fi
  \parbox{\indentedwidth}{\raggedright
       \ignorespaces #1\\[2\medskipamount]%
       \ifx\@empty\fromsig
           \fromname
       \else \fromsig \fi\strut}%
   \par}
\makeatother
% Some of the article customisations are relevant for this class

\signature{Justin Katz \\
Mathematics Department\\
Purdue University\\
1610 Purdue Mall\\
West Lafayette, IN 47907\\	(219) 210-9452 \\jukatz93@gmail.com\\
} % Goes after the closing (ie at the end of the letter, with space for a signature)
\address{640 Ferry Street \\ Lafayette, Indiana  47901 \\ USA} 
\date{}
% Alternatively, these may be set on an individual basis within each letter environment.

\begin{document}
\begin{letter}{}

\opening{To whom it may concern,} % eg Hello.
I am writing to apply to be a Hans J. Zassenhaus Visiting Assistant professor or Arnold Ross Visiting Assistant professor at the Ohio State university. 

I recently had an opportunity to speak in the geometry and topology seminar at the Ohio State University. During my visit, I had an opportunity to meet with professors J. LaFont and M. Lipnowski, with whom my research interests are aligned.  I would cherish an opportunity to work with these professors, and many others at OSU. 

I am a Ph.D. candidate at Purdue University, under the direction of D.B. McReynolds. I expect to complete my degree by May of 2023. My research interests include geometry and topology, spectral theory, number theory, representation theory, harmonic analysis, and interactions thereof. My thesis work concerns an absolute spectral rigidity result, which applies to a large class of arithmetic hyperbolic 2 and 3 manifolds.  One application of this theorem proves that all principal congruence Hurwitz surfaces are spectrally rigid. I have also co-authored a paper in which we construct large collections of algebraic and differential geometric objects which are indistinguishable by means of various invariants. I consider teaching to be an essential role of the mathematician, and am committed to fostering a safe, inclusive, and open environment for students.  

Please find my curriculum vitae and research and teaching statements attached.  Please contact me for any further information. 



\closing{Sincerely,} % eg Regards,


\end{letter}
\end{document}
