\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{amsthm}
\usepackage{hyperref}
\input{mydefs}

\title{Homogeneous Spaces as Quotients of Groups}
\author{Justin Katz}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle
This statement and proof draws heavily on the exposition of Paul Garrett, in the appendix of \url{http://www.math.umn.edu/~garrett/m/mfms/notes/02_solenoids.pdf}.

Let $X$ be a locally compact Hausdorff space and $G$ be a topological group acting continuously, transitively on $X$. Fix a point $x\in X$ and let $G_x$ be the isotropy subgroup of $G$ at $x$. 
\begin{claim}
	The $G$-\emph{space} $X$ is homeomorphic to the \emph{quotient space} $G/G_x$ under the assignment
	\begin{equation*}
		gG_x \mapsto gx
	\end{equation*}
\end{claim}
\begin{proof}
	By the transitivity of the $G$-action on $X$, the map $gG_x\mapsto gx$ surjects. Because $G_x$ fixes $x$, the map injects. To prove the claim, it suffices to show that the map is continuous and open.

The topology on the quotient with projection $\pi:G\to G/G_x$ is uniquely characterized by the condition that any continuous map out of $G$ that is constant on $G_x$ factors uniquely through $\pi$ to a continuous map out of $G/G_x$. The map $g \mapsto gx$ is continuous as a restriction of the action, and is constant on $G_x$ by definition of isotropy. Thus $g \mapsto gx$ factors uniquely through $\pi$ to a continuous map out of $G/G_x$. The map $gG_x \mapsto gx$ fits the bill, so is continuous.

To prove that $gG_x\mapsto gx$ is open, let $U$ be a neighborhood of $g \in G$. For reasons that will become apparent later, we want a compact neighborhood $V$ of $1$ so that $gV^2=\{gvh:v,h\in V\}\subset U$. To show such a compact set $V$ exists first show the result at $g=1$.The inverse image of the open $U$ under the (continuous) product map $h\times k\mapsto hk$ is again open. Open sets in the product topology are generated by products of opens in the producands, so the inverse image of $U$ under multiplication contains a product of opens $W_1\times W_2$ each containing $1$. Let $W=W_1\cap W_2$ so that $W^2\subset W_1 \cdot W_2 \subset U$ where the last containment comes from the definition of $W_1\times W_2$ as a subset of the inverse image of $U$ under multiplication. Furthermore $G$ is Hausdorff, so there is some neighborhood $W'$ of $1$ contained in $W$ such that $\overline{W'}$ is compact and sits inside $W$. Let $V=\overline{W'}$ so that $V^2\subset W^2\subset U$. For generic $g$ with neighborhood $U$, the open $g^{-1}U$ is a neighborhood of $1$, and the above discussion gives the result. We can \emph{balance} $V$ about $1$ (i.e. make it such that $V=V^{-1}$) by setting $V\mapsto V \cap V^{-1}$.

Next, we show that the whole group $G$ can be covered by countably many translates of the compact $V$. First, we show the result for some open $W$ in $V$. Let $\{U_1,U_2,...\}$ be a (countable) basis for $G$. For each $g\in G$, by the definition of basis, the open $gW$ is the union of those $U_i\subset gW$. As such, for each $g\in G$ there is a smallest index index $j(g)$ such that $g \in U_{j(g)} \subset gW$. For each index $i$ pick some $g_i$ in $j^{-1}(i)$ so that $g_i \in U_i \subset g_i W$. By definition of the map $g\mapsto j(g)$, we have $j^{-1}(i)\subset U_i \subset g_i W$. Taking the union over all (countably many) indices $i$, $\cup j^{-1}(i)=G \subset \cup g_i W$ as desired. We can certainly replace $W$ by its compact superset $V$ so that $G=\cup g_i V$ as claimed.

We are now ready to prove that the map $gG_x\mapsto gx$ is open. Recall that $U$ is a neighborhood of some point $g\in G$, $V$ is a balanced compact in $U$ such that $V^2\subset U$. We want to show that $Ux$ is open.  Recall the version of the \emph{Baire category theorem}:
\begin{quote}
	\emph{A locally compact Hausdorff space is not a countable union of nowhere dense sets}
\end{quote}
In particular, by transitivity of the group action we can cover the space $X$ by countably many $Vx$ translates $X=\cup g_i V x$. Note that each translate $g_i V x$ is closed, being the continuous image of a compact $g_i V$ in a Hausdorff space.  By Baire, some $g_m V x$ contains a nonempty open $S$. Let $h$ be such that $g_m h x \in S$ and write 
	\begin{equation*}
		gx=g(g_m h)^{-1}(g_m h)x\in g h^{-1}g_{m}^{-1} S
	\end{equation*} 
The rightmost set in the above display is again open in $X$ because translation in $X$ by a fixed element of $G$ is a homeomorphism. Compute
	\begin{align*}
		gx \in g h^{-1}g_{m}^{-1} S  & \subset gh^{-1}g_{m}^{-1} g_m V x  &\qquad \left(\text{By definition of S} \right) \\
						& \subset g h^{-1} V x  \\
						& \subset g V^{-1} V x \\
						& = g V^2 x \qquad &\left( V \text{ is balanced about } 1 \right) \\
						& \subset Ux  \qquad &\left( \text{By definition of } V \right)	,			
	\end{align*}
meaning $gx$ is an interior point of $Ux$. The group element $g \in U$ was arbitrary so $Ux$ is open, proving the claim.
\end{proof}
\begin{remark}
If\footnote{I essentially follow Warner in his text \emph{Foundations of Differntiable Manifolds and Lie Groups}, roughly page 120} $X$ is a smooth manifold and $G$ is a Lie group acting on $X$ smoothly, then the homeomorphism in the conclusion of the above claim is actually a diffeomorphism. Indeed as the isotropy subgroup $G_x$ is closed, the quotient $G/G_x$ has a unique smooth structure so that any smooth map out of $G$ constant on $G_x$ factors uniquely through the projection $\pi$ to a smooth map out of $G/G_x$. Because $G/G_x$ is already homeomorphic to $G/G_x$ and (by the mapping property of quotients) the map $f:gG_x\mapsto gx$ is smooth, (by the inverse function theorem) it suffices to show that the differential  $df_{1G_x}:T(G/G_x)_{1G_x}\to TX_x$ is nonsingular. Note that the map $h:G\to X$ defined by $g\mapsto gx$ is the composition $f\circ \pi$. Thus, to show that $df_{1G_x}$ is nonsingular, it suffices to show that the kernel of $dh_1$ is exactly the kernel of $d\pi_1$, i.e. the tangent space $T(G_x)_{1}$. One direction is easy: $\ker dh_1$ certainly contains $T(G_x)_1$, because $h$ is constant on $G_x$. To prove the other direction, let $z \in \ker dh_1$ and let $Z$ be the corresponding left invariant vector field on $G$. Recall that the left invariance of $Z$ is the equality $d(L_\gamma)_{Z(\cdot)}= Z\circ L_\gamma(\cdot)$ where $L_\gamma :g\mapsto \gamma g$ is the (smooth) left action of $G$ on itself. That $Z$ corresponds to $z$ means that $Z$ is the unique vector field such that $\frac{d}{dr}\exp(rZ)|_0=z$. To show $z\in T(G_x)_1$ it suffices to show that $\exp(tZ) \in G_x$ for all $t\in \R$, meaning $\exp(tZ)$ fixes $x$ for all $t$. Consider the curve $\alpha: t\mapsto h(\exp(tZ))$ in $M$. If the tangent vector to $\alpha$ is zero at every $t$ then $\alpha$ is constant. In particular, $\alpha(0)=h(1)=x$ so if $\alpha$ is constant then $\exp(tZ)$ fixes $x$ for all $t$ and is thus in $G_x$. To prove that the tangent vector to $\alpha$ is zero, first compute for $t=0$
	\begin{align*}
		d(\alpha)_0&=d(h)_{1}\circ \frac{d}{dr}\exp(r Z)|_0\\
				 &=dh_1(z) &\qquad \left(Z \text{ corresponds to } z\right)\\
				 &=0		    &\qquad \left(z\in \ker{d(h)_1} \right).
	\end{align*}
To prove that $\frac{d}{dr}\alpha(r)|_t=0$ for all $t$ notice that the map $h$ is invariant under conjugation by a group element $\gamma$ i.e. $\gamma \cdot h \circ L_\gamma^{-1} (g)=\gamma \cdot \gamma^{-1} \cdot g x=gx=h(g)$. In particular, for $\gamma=\exp(tZ)$ compute
	\begin{align*}
		\frac{d}{dr}\alpha(r)|_t &= d(h)_{\exp(tZ)} \circ \frac{d}{dr} \exp(r Z)|_t &\left( \text{Chain rule} \right) \\
				  &= d(\exp(tZ)\cdot h \circ L_{\exp(-tZ)})_{\exp(tZ)}\circ \frac{d}{dr}e^{r Z})|_t  &\left( \text{Invariance of $h$ under conjugation} \right)\\
				  &= d(\exp(tZ) \cdot h )_{L_{\exp(-tZ)}(\exp(tZ))}\circ \frac{d}{dr}L_{\exp(-tZ)} \exp(r Z)|_t &\left( \text{Chain rule} \right) \\
				  &=d(\exp(tZ) \cdot h )_1 \circ \frac{d}{dr} \exp(r Z)|_0 &\left( \text{Definition of left action, changing variables $t\mapsto r-t$} \right) \\
				  &=0 &\left(\frac{d}{dr} \exp(r Z)|_0=z\in \ker dh_1\right).
	\end{align*}
	Thus the curve $\alpha$ is constant, so $\exp(tZ)$ fixes $x$ for all $t$, meaning the tangent vector $z$ corresponding to $Z$ is in $T_1 G_x$. Therefore $\ker dh_1 = T_1 G_x$, so that smooth bijection $f:gG_x \mapsto gx$ has nonsingular derivative, and is thus a diffeomorphism as desired.
\end{remark}
\end{document}  